\documentclass[10pt,a4paper]{article}
\usepackage[margin=2.0cm]{geometry}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{array}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage[colorlinks=true]{hyperref}

\definecolor{fpgablue}{RGB}{13,71,161}
\definecolor{okgreen}{RGB}{27,94,32}
\definecolor{badred}{RGB}{183,28,28}
\definecolor{rulegray}{RGB}{120,120,120}
\hypersetup{linkcolor=fpgablue, pdftitle={FPGA Tick-to-Trade Implementation Results}}

\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt}

\newcommand{\fail}{\textcolor{badred}{\textbf{FAIL}}}
\newcommand{\pass}{\textcolor{okgreen}{\textbf{PASS}}}

\begin{document}

{\LARGE\bfseries\color{fpgablue} FPGA Tick-to-Trade --- Implementation Results}\hfill
{\small\color{rulegray} June 2026}\\[-1pt]
{\large Synthesis \& place-and-route results, with a baseline for tracking optimisation.}

\textcolor{rulegray}{\rule{\linewidth}{1pt}}

This document records real Vivado synthesis + implementation (placed \& routed)
results for \texttt{tick\_to\_trade\_top}, so successive optimisations can be
compared. Reports live in \texttt{results/synthesis/} and
\texttt{results/implementation/}.

\section*{Run B0 --- Baseline (registered order-book array)}

\begin{center}
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{@{}l l@{}}
\toprule
\textbf{Item} & \textbf{Value} \\
\midrule
Target device       & \texttt{xcu200-fsgd2104-2-e} (Alveo U200 / UltraScale+ VU9P --- AWS F1 class) \\
Tool                & Vivado 2025.x, out-of-context (no physical pin constraints) \\
Target clock        & 200\,MHz (5.000\,ns period) \\
Flow                & synth\_design $\rightarrow$ opt $\rightarrow$ place $\rightarrow$ route (fully routed) \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Timing --- \fail\ at 200\,MHz}
\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}l r@{}}
\toprule
\textbf{Metric} & \textbf{Value} \\
\midrule
Worst Negative Slack (WNS)  & \textcolor{badred}{$-2.373$\,ns} \\
Failing endpoints           & 41{,}206 \\
Total Negative Slack (TNS)  & $-51{,}387$\,ns \\
Max achievable clock        & $\approx$\,135\,MHz \;($1/(5.0{+}2.373)$\,ns) \\
Implied tick-to-trade       & 39 cycles $\times$ 7.37\,ns $\approx$ \textbf{290\,ns} (not 195\,ns) \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Utilisation (placed) --- tiny footprint}
\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}l r r r@{}}
\toprule
\textbf{Resource} & \textbf{Used} & \textbf{Available} & \textbf{\%} \\
\midrule
CLB LUTs        & 36{,}057 & 1{,}182{,}240 & 3.05 \\
CLB Registers   & 37{,}672 & 2{,}364{,}480 & 1.59 \\
CARRY8          &    265   &   147{,}780  & 0.18 \\
Block RAM Tile  &      \textbf{0} &     2{,}160 & 0.00 \\
DSPs            &      \textbf{0} &     6{,}840 & 0.00 \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Power \& DRC}
\begin{center}
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{@{}l l@{}}
\toprule
Total on-chip power & 2.844\,W (dynamic 0.371\,W, static 2.473\,W; low confidence) \\
Routing             & fully routed (52{,}563 / 52{,}563 routable nets) \\
DRC                 & 2 critical warnings: NSTD-1, UCIO-1 (unconstrained I/O --- expected \\
                    & \quad in OOC flow with no pin constraints; benign for F1-shell integration) \\
\bottomrule
\end{tabular}
\end{center}

\subsection*{Critical path \& root cause}
\begin{center}
\small
\begin{tabular}{@{}l l@{}}
\toprule
Source       & \texttt{u\_book/entries\_reg[7][shares][8]} \\
Destination  & \texttt{u\_book/entries\_reg[219][shares][18]} \\
Data delay   & 7.036\,ns --- \textbf{route 5.969\,ns (85\%)}, logic 1.067\,ns, 11 logic levels \\
\bottomrule
\end{tabular}
\end{center}

\noindent\textbf{Root cause:} the order-book \texttt{entries[]} array (256 $\times$
130 bits) synthesised to \textbf{37\,K flip-flops (BRAM = 0)} instead of memory,
and the RESCAN FSM reads \emph{all} entries combinationally. The huge fan-out of
that register file makes the path \textbf{route-dominated (85\%)}, so it misses
200\,MHz badly. Logic is fine ($<$1.1\,ns); the problem is physical (FF array +
combinational scan).

\section*{Planned fix --- Run B1 (BRAM-based order book)}

Convert \texttt{entries[]} to a true \textbf{BRAM with synchronous (registered)
reads}. The RESCAN FSM already walks one entry per cycle, so a 1-cycle BRAM read
drops in naturally; the Add fast-path stays O(1) (writes + level-insert from the
message, no read). This removes both the 37\,K-FF array and the route blow-up and
should close 200\,MHz.

\textcolor{rulegray}{\rule{\linewidth}{1pt}}

\section*{Correction 1 --- Run B1 (BRAM order book, measured)}

The BRAM order book was implemented and re-run through synth + implementation on
the VU9P. \textbf{Outcome: the resource/route problem is fixed, but timing still
fails --- for a new reason.}

\begin{center}
\renewcommand{\arraystretch}{1.25}
\begin{tabular}{@{}l c c c@{}}
\toprule
\textbf{Metric} & \textbf{B0 (FF array)} & \textbf{B1 (BRAM)} & \textbf{Change} \\
\midrule
CLB Registers (FF)      & 37{,}672           & \textbf{4{,}308}   & $-89\%$ \;\checkmark \\
CLB LUTs                & 36{,}057           & \textbf{6{,}790}   & $-81\%$ \;\checkmark \\
Block RAM Tile          & 0                  & \textbf{2}         & inferred \;\checkmark \\
DSPs                    & 0                  & 0                  & --- \\
Failing endpoints       & 41{,}206           & \textbf{3{,}612}   & $-91\%$ \;\checkmark \\
Total Negative Slack    & $-51{,}387$\,ns    & \textbf{$-2{,}522$\,ns} & $-95\%$ \;\checkmark \\
Worst Negative Slack    & $-2.373$\,ns       & \textcolor{badred}{$-4.563$\,ns} & worse \\
Total on-chip power     & 2.844\,W           & 2.607\,W           & $-8\%$ \\
Timing @ 200\,MHz       & \fail              & \fail              & not yet closed \\
\bottomrule
\end{tabular}
\end{center}

\noindent\textbf{What worked:} \texttt{entries[]} now infers block RAM (2 tiles)
instead of 37\,K flip-flops; LUTs/FFs collapsed $\sim$85\%; the route-dominated
FF-array path is gone --- failing endpoints and total violation both dropped
$\sim$90--95\%. The original root cause is resolved.

\noindent\textbf{New critical path (B1):}
\begin{center}
\small
\begin{tabular}{@{}l l@{}}
\toprule
Source       & \texttt{u\_book/bid\_lv\_reg[cnt][0]} (level cache) \\
Destination  & \texttt{risk\_reason[1]} (output pin / OBUF) \\
Data delay   & 5.168\,ns --- logic 2.494\,ns (48\%), route 2.673\,ns (52\%), 14 levels, 5$\times$CARRY8 \\
\bottomrule
\end{tabular}
\end{center}

\noindent\textbf{Why WNS got worse:} this is no longer a routing problem --- it is
a long \emph{combinational} chain: book level cache $\rightarrow$ strategy
(depth-weighted cross-multiply) $\rightarrow$ \texttt{risk\_check} (price-collar
subtract, position add) $\rightarrow$ straight to the decision/\texttt{risk\_reason}
output pins, with no register in between. \texttt{risk\_check} was added as a
\emph{zero-latency combinational} gate (to preserve 195\,ns); once the FF-array
path was removed, this chain became the worst path.

\section*{Correction --- Run B2 (register the risk/decision stage), measured}

Made \texttt{risk\_check} a \textbf{1-cycle pipeline} (registered outputs) and
registered the decision output ports, splitting
book\,$\rightarrow$\,strategy\,$\rightarrow$\,risk\,$\rightarrow$\,pin. Cost:
\textbf{+1 cycle} (39\,$\rightarrow$\,40 cycles, 195\,$\rightarrow$\,\textbf{200\,ns}).
\texttt{tick\_to\_trade\_top} was re-synthesised + implemented on the VU9P:

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}l c c c@{}}
\toprule
\textbf{Metric} & \textbf{B1 (comb.\ risk)} & \textbf{B2 (registered)} & \textbf{Change} \\
\midrule
WNS                  & $-4.563$\,ns & \textbf{$-2.217$\,ns} & \textbf{+2.35\,ns} \;\checkmark \\
Failing endpoints    & 3{,}612      & \textbf{87}           & $-$3{,}525 (98\% fewer) \\
Worst-path dest.     & \texttt{risk\_reason} & \texttt{s\_axil\_awready} & moved off the trade path \\
Worst-path logic     & 14 levels    & \textbf{2 levels} (LUT2+OBUF) & --- \\
\bottomrule
\end{tabular}
\end{center}

\noindent\textbf{B2 closed the trading datapath.} The original decision cone
(book\,$\rightarrow$\,strategy\,$\rightarrow$\,risk\,$\rightarrow$\,pin) no longer
appears anywhere in the failing list. The residual $-2.217$\,ns is no longer a
logic problem --- it is a \textbf{clock-insertion + I/O-budget artifact of the
out-of-context run}:

\begin{itemize}\itemsep2pt
  \item The worst path (\texttt{s\_axil\_bvalid\_reg}\,$\rightarrow$\,\texttt{s\_axil\_awready},
  the AXI-Lite latency-readout port) has only \textbf{2 logic levels} but rides
  a \textbf{2.714\,ns clock-insertion delay} --- one global BUFG fanning out to
  4{,}757 loads, i.e.\ 54\% of the 5\,ns period gone before any logic.
  \item Insertion delay cancels on internal register\,$\rightarrow$\,register
  paths (which meet 200\,MHz) but is charged in full against I/O paths. With the
  former 40\% (2.0\,ns) I/O budget the boundary was unsatisfiable even though the
  logic is fine.
\end{itemize}

\noindent\textbf{Fix (constraints, not RTL):} in \texttt{rtl/top.xdc} ---
(1) \texttt{set\_false\_path} the whole \texttt{s\_axil\_*} boundary (a slow,
host-polled status port, not a 200\,MHz datapath; the Shell closes its timing
in-context); (2) right-size the boundary I/O budget 40\%\,$\rightarrow$\,10\%
(0.5\,ns) to reflect that the real F1 build places the clock in clock regions
(insertion $\sim$0.3\,ns) and the Shell registers the CL/SH interface. Re-run
synth+impl to confirm; the internal logic already meets 200\,MHz.

\textcolor{rulegray}{\rule{\linewidth}{1pt}}

\section*{Correction 2 --- Multi-symbol engine ($\times$4), measured}

\texttt{multi\_symbol\_top} (NSYMBOLS=4: four per-symbol books routed by
\texttt{stock\_locate}) was synthesised + implemented on the VU9P.

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}l c c c@{}}
\toprule
\textbf{Metric} & \textbf{single-sym B1} & \textbf{multi-sym ($\times$4)} & \textbf{Note} \\
\midrule
Block RAM Tile       & 2        & \textbf{8}      & exactly 4$\times$ --- one BRAM book/symbol \;\checkmark \\
CLB LUTs             & 6{,}790  & 19{,}942        & $\sim$3$\times$ (shared strategy/risk) \\
CLB Registers (FF)   & 4{,}308  & 8{,}984         & $\sim$2$\times$ \\
WNS                  & $-4.563$\,ns & \textcolor{badred}{$-5.362$\,ns} & worse \\
Failing endpoints    & 3{,}612  & 10{,}899        & --- \\
Power                & 2.607\,W & 2.869\,W        & --- \\
Routing              & clean    & fully routed, 0 errors & \checkmark \\
\bottomrule
\end{tabular}
\end{center}

\noindent\textbf{Confirms the BRAM book scales:} 4 symbols $\rightarrow$ 8 BRAM
tiles (4$\times$2), no FF explosion. \textbf{Worst path} (\texttt{gbook[3] bid\_lv
$\rightarrow$ m\_axis\_tvalid}, 18 levels, 8$\times$CARRY8) is the same
combinational book\,$\rightarrow$\,strategy\,$\rightarrow$\,risk\,$\rightarrow$\,pin
chain, \emph{lengthened} by the 4-way active-symbol output mux. This run pre-dates
the register fix; the fix has since been applied to \texttt{multi\_symbol\_top}
too (re-run pending).

\textcolor{rulegray}{\rule{\linewidth}{1pt}}

\section*{Correction 3 --- Run B3 (pipeline strategy + collar reference)}

Analysis of the routed B1/M1 reports showed that registering only the output
(B2) is necessary but \emph{may not be sufficient}: two structural problems
remain on the decision cone.

\begin{center}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}p{0.46\linewidth} p{0.46\linewidth}@{}}
\toprule
\textbf{Residual issue (from B1/M1 routed paths)} & \textbf{Why it still hurts after B2} \\
\midrule
Book\,$\rightarrow$\,\texttt{mid\_price} (add+shift)\,$\rightarrow$ \texttt{risk\_check} collar\,$\rightarrow$\,output. & The collar reference was a live combinational function of the book best price, feeding the risk comparators. \\
Strategy $\Sigma$(weight$\cdot$size) \emph{and} the 64-bit cross-multiply in one cycle. & Weighted-sum + cross-multiply + compare were one deep cone (route 52--58\%). \\
\bottomrule
\end{tabular}
\end{center}

\noindent\textbf{Fix (B3):}
\begin{itemize}\itemsep2pt
  \item \texttt{strategy\_imbalance} is now a \textbf{2-stage pipeline}: stage~1
  registers the depth-weighted volumes $w_{bid}/w_{ask}$ and the book snapshot;
  stage~2 does the cross-multiply + lot sizing + fire decision. This isolates the
  wide summation from the cross-multiply compare.
  \item Both tops register the collar reference (book mid) \textbf{2 deep}
  (\texttt{mid\_r1}/\texttt{mid\_r2}) so the risk comparators start at a flip-flop,
  not the \texttt{book\,$\rightarrow$\,mid} adder, and stay aligned with the now
  2-cycle strategy. (Bonus: this also removes a 1-cycle order/mid skew that
  existed in B2.)
\end{itemize}

\noindent\textbf{Cost:} \textbf{+1 cycle} (tick-to-trade 40\,$\rightarrow$\,41
cycles, 200\,$\rightarrow$\,\textbf{205\,ns}). The split halves the longest
combinational segment on both the strategy and risk sub-paths.

\noindent\textbf{Status:} implemented and verified in simulation ---
\textbf{95/95 cocotb tests pass}, lint clean, tick-to-trade measured at
\textbf{205\,ns}, multi-symbol decisions still correctly attributed. \emph{Not
yet re-synthesised} --- re-run synth+impl on \texttt{tick\_to\_trade\_top}
(then \texttt{multi\_symbol\_top}) to populate the real B3 timing and confirm
200\,MHz closure.

\textcolor{rulegray}{\rule{\linewidth}{1pt}}

\section*{Correction 4 --- Run B4 (XDC: false-path AXI-Lite + 10\% I/O budget), measured}

Re-synthesised + implemented \texttt{tick\_to\_trade\_top} with the B4 constraints
fix. The AXI-Lite false-path + 10\% I/O budget corrected most of the B2 I/O
violations:

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}l c c c@{}}
\toprule
\textbf{Metric} & \textbf{B2} & \textbf{B4} & \textbf{Change} \\
\midrule
WNS                  & $-2.217$\,ns & \textbf{$-0.578$\,ns} & \textbf{+1.64\,ns} \;\checkmark \\
Failing endpoints    & 87           & \textbf{47}           & $-$40 \\
Worst-path logic     & 2 levels (AXI-Lite) & \textbf{1--2 levels (FF$\to$OBUF)} & all FF-to-output pins \\
Internal reg$\to$reg & meets 200\,MHz & meets 200\,MHz & \textbf{confirmed} \\
\bottomrule
\end{tabular}
\end{center}

\noindent\textbf{Remaining 47 violations are all FF$\to$OBUF paths with 0--1
logic levels:}

\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}l l c r@{}}
\toprule
\textbf{Destination (output pin)} & \textbf{Source (FF)} & \textbf{Levels} & \textbf{Slack} \\
\midrule
\texttt{s\_axis\_tready}   & \texttt{u\_book/state\_reg[0]} & 2 (LUT3+OBUF) & $-0.578$\,ns \\
\texttt{risk\_reason[0]}   & \texttt{risk\_reason\_reg[0]}  & 1 (OBUF)      & $-0.461$\,ns \\
\texttt{m\_axis\_tdata[*]} & \texttt{m\_axis\_tdata\_reg[*]}& 1 (OBUF)      & $\sim$$-0.44$\,ns \\
\bottomrule
\end{tabular}
\end{center}

\noindent\textbf{Root cause (unchanged):} OOC BUFG clock insertion 2.658\,ns
across fanout $\sim$4,752 loads. This delay is charged in full on FF$\to$output-pin
paths even after the I/O budget reduction, because the OBUF propagation
($\sim$1.1\,ns) + insertion (2.658\,ns) $\approx$ 3.76\,ns exceeds the 0.5\,ns
required budget. In the real F1 build the Shell registers every CL/SH boundary
port and the clock is placed in clock regions (insertion $\sim$0.3\,ns); these
paths do not exist in-context.

\noindent\textbf{Fix (B5):} false-path \emph{all} remaining CL/SH boundary
ports (\texttt{s\_axis\_*}, \texttt{m\_axis\_*}, \texttt{risk\_*},
\texttt{halt}) so STA evaluates only internal register$\to$register paths, which
already meet 200\,MHz. The \texttt{top.xdc} has been updated accordingly.

\section*{Run history}
\begin{center}
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{@{}l l c l@{}}
\toprule
\textbf{Run} & \textbf{Change} & \textbf{WNS} & \textbf{Status} \\
\midrule
B0 & registered FF order-book array          & $-2.373$\,ns & \fail\ (route-bound) \\
B1 & \texttt{entries[]} $\to$ BRAM           & $-4.563$\,ns & \fail\ (combinational chain) \\
M1 & multi-symbol $\times$4 (BRAM, comb.\ risk) & $-5.362$\,ns & \fail\ (chain + sym mux) \\
B2 & register risk/decision stage            & $-2.217$\,ns & \fail\ I/O only --- \textbf{datapath closed}; 87 endpoints \\
B3 & pipeline strategy (2-stage) + collar ref & TBD          & RTL done (205\,ns sim, 95/95); \emph{re-run synth} \\
B4 & XDC: false-path AXI-Lite + 10\% I/O budget & $-0.578$\,ns & \fail\ FF$\to$OBUF only (47 eps); internal logic \textbf{MEETS} 200\,MHz \\
B5 & XDC: false-path all CL/SH boundary ports  & TBD          & constraints done; \emph{re-run synth} \\
\bottomrule
\end{tabular}
\end{center}

\vfill
\textcolor{rulegray}{\rule{\linewidth}{0.5pt}}
{\footnotesize\color{rulegray}
Source reports: \texttt{results/implementation/*.rpt}, \texttt{results/synthesis/*.rpt}.
Re-run with \texttt{vivado -mode batch -source build/synth.tcl} on a $\geq$24\,GB machine.}

\end{document}
